---
title: "CCR Journal Health Report"
author: "Editorial team"
date: today
format:
  html:
    toc: true
    df-print: kable
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(DT)

# .env holds OJS_BASE_URL and OJS_ADMIN_TOKEN; ojs.R defaults to OJS_TOKEN.
readRenviron(".env")
Sys.setenv(OJS_TOKEN = Sys.getenv("OJS_ADMIN_TOKEN"))

source("ojs.R")

bucket <- function(label) {
    l <- tolower(label %||% "")
    dplyr::case_when(
        grepl("publish|production|copyedit|schedul", l) ~ "Accepted",
        grepl("declin|reject", l) ~ "Rejected",
        TRUE ~ "Pending"
    )
}

parse_dt <- function(x) as.POSIXct(x, tz = "UTC", format = "%Y-%m-%d %H:%M:%S")
```

```{r load-data}
#| cache: true
client <- ojs_client()
submissions <- ojs_submissions(client)
decisions <- ojs_decisions_bulk(client, submissions$id)
```

::: {.panel-tabset}

## Submissions and reviewing

### Submissions by year and final status

Pipeline volume per year and what happened to each cohort.

```{r submissions-by-year-status}
subs_by_section <- submissions |>
    mutate(
        year    = substr(date_submitted, 1, 4),
        status  = bucket(status_label),
        section = ifelse(section_id == 122L,
                         "Research Article",
                         "Other (special issues, notes, tools)")
    ) |>
    filter(nzchar(year))

build_year_status <- function(df, caption) {
    df |>
        mutate(status = factor(status,
                               levels = c("Accepted", "Rejected", "Pending"))) |>
        count(year, status, .drop = FALSE) |>
        pivot_wider(names_from = status, values_from = n, values_fill = 0) |>
        mutate(
            Total = Accepted + Rejected + Pending,
            `Accept rate` = ifelse(
                (Accepted + Rejected) > 0,
                sprintf("%.0f%%", 100 * Accepted / (Accepted + Rejected)),
                "—"
            )
        ) |>
        arrange(year) |>
        select(Year = year, Accepted, Rejected, Pending, Total, `Accept rate`) |>
        kable(caption = caption)
}

subs_by_section |>
    filter(section == "Research Article") |>
    build_year_status("Research Article submissions per year by final status")
```

```{r submissions-by-year-status-other}
subs_by_section |>
    filter(section != "Research Article") |>
    build_year_status(
        "Other (special issues, notes, tools) submissions per year by final status"
    )
```

::: {.callout-note collapse="true" title="Column definitions and section split"}
- **Year** — year of `dateSubmitted`.
- **Accepted** — published or in copyediting/production.
- **Rejected** — declined (desk-reject or post-review).
- **Pending** — still in queue, review, or revision.
- **Accept rate** — Accepted / (Accepted + Rejected). Excludes Pending papers
  from the denominator. Recent submission-years still have many Pending papers
  whose outcomes will move the rate as they finalize.

The split is by OJS `sectionId` of the publication's first version. Section
122 is treated as "Research Article" (the main section, used every year since
2018); all other sections are bundled as "Other" (special issues, research
notes, tool announcements — the API doesn't expose section names, so they
can't be split further from this report alone).
:::

### Decision turnaround times

How fast papers move from submission to first decision, and through subsequent
review rounds to a final outcome.

```{r decision-times-overall}
dec <- decisions |>
    mutate(date_decided = parse_dt(date_decided)) |>
    filter(!is.na(date_decided))

subs_dt <- submissions |>
    mutate(
        date_submitted = parse_dt(date_submitted),
        final          = bucket(status_label)
    ) |>
    select(id, date_submitted, final)

final_decisions <- dec |>
    mutate(kind = dplyr::case_when(
        grepl("accept", tolower(label)) & !grepl("production", tolower(label)) ~ "accept",
        grepl("declin", tolower(label)) ~ "decline",
        TRUE ~ NA_character_
    )) |>
    filter(!is.na(kind)) |>
    group_by(submission_id, kind) |>
    summarise(final_date = max(date_decided), .groups = "drop") |>
    pivot_wider(
        names_from = kind, values_from = final_date,
        names_prefix = "date_"
    )

per_sub <- dec |>
    group_by(submission_id) |>
    summarise(
        first_decision = min(date_decided),
        had_review = any(!is.na(review_round_id) & review_round_id > 0),
        .groups = "drop"
    ) |>
    inner_join(subs_dt, by = c("submission_id" = "id")) |>
    left_join(final_decisions, by = "submission_id") |>
    mutate(
        final_decision_date = dplyr::case_when(
            final == "Accepted" ~ date_accept,
            final == "Rejected" ~ date_decline,
            TRUE ~ as.POSIXct(NA)
        ),
        days_to_first = as.numeric(difftime(first_decision, date_submitted, units = "days")),
        days_to_last = as.numeric(difftime(final_decision_date, date_submitted, units = "days"))
    )

per_sub <- per_sub |>
    mutate(year = substr(as.character(date_submitted), 1, 4)) |>
    filter(year >= "2024")

summary_row <- function(df) {
    final_df <- df |>
        filter(final != "Pending", had_review, !is.na(final_decision_date))
    tibble(
        first_N      = sum(!is.na(df$days_to_first)),
        first_median = round(median(df$days_to_first, na.rm = TRUE), 1),
        first_mean   = round(mean(df$days_to_first, na.rm = TRUE), 1),
        final_N      = nrow(final_df),
        final_median = round(median(final_df$days_to_last, na.rm = TRUE), 1),
        final_mean   = round(mean(final_df$days_to_last, na.rm = TRUE), 1)
    )
}

by_year <- per_sub |>
    filter(nzchar(year)) |>
    group_by(Year = year) |>
    group_modify(~ summary_row(.x)) |>
    ungroup() |>
    arrange(Year)

total <- summary_row(per_sub) |>
    mutate(Year = "Total") |>
    select(Year, everything())

bind_rows(by_year, total) |>
    setNames(c("Year", "N", "Median", "Mean", "N", "Median", "Mean")) |>
    kable(caption = "Time to first and final decision, per submission year") |>
    add_header_above(c(
        " " = 1,
        "First decision (days)" = 3,
        "Final decision (days)" = 3
    ))
```

Review-round duration: how long each round of revision-and-review takes.

```{r decision-times-rounds}
# A round closes at the *first* AE/EiC decision that resolves the review
# (Accept / Decline / Request Revisions / Resubmit). Later same-round_id
# decisions are admin events ("New Review Round" marking author resubmission)
# or post-revision accepts that no longer involve reviewers — including those
# in round_end would inflate "deciding" time. See submission 8874 for the
# canonical example.
round_closing_labels <- c(
    "Accept Submission", "Decline Submission",
    "Request Revisions", "Resubmit for Review"
)

rounds <- dec |>
    filter(!is.na(review_round_id), review_round_id > 0,
           label %in% round_closing_labels) |>
    group_by(submission_id, review_round_id) |>
    summarise(round_end = min(date_decided), .groups = "drop") |>
    arrange(submission_id, round_end) |>
    group_by(submission_id) |>
    mutate(
        round_num = row_number(),
        prev_end = lag(round_end)
    ) |>
    ungroup() |>
    inner_join(subs_dt |> select(id, date_submitted),
        by = c("submission_id" = "id")
    ) |>
    mutate(
        round_start = coalesce(prev_end, date_submitted),
        round_days  = as.numeric(difftime(round_end, round_start, units = "days")),
        year        = substr(as.character(date_submitted), 1, 4)
    ) |>
    filter(year >= "2024")

round_summary_row <- function(df) {
    r1 <- df |> filter(round_num == 1)
    rN <- df |> filter(round_num >= 2)
    per_sub_rounds <- df |>
        group_by(submission_id) |>
        summarise(n_rounds = max(round_num), .groups = "drop")
    tibble(
        r1_N       = nrow(r1),
        r1_median  = round(median(r1$round_days, na.rm = TRUE), 1),
        r1_mean    = round(mean(r1$round_days, na.rm = TRUE), 1),
        rN_N       = nrow(rN),
        rN_median  = round(median(rN$round_days, na.rm = TRUE), 1),
        rN_mean    = round(mean(rN$round_days, na.rm = TRUE), 1),
        avg_rounds = round(mean(per_sub_rounds$n_rounds), 2)
    )
}

by_year_rounds <- rounds |>
    group_by(Year = year) |>
    group_modify(~ round_summary_row(.x)) |>
    ungroup() |>
    arrange(Year)

total_rounds <- round_summary_row(rounds) |>
    mutate(Year = "Total") |>
    select(Year, everything())

bind_rows(by_year_rounds, total_rounds) |>
    setNames(c("Year", "N", "Median", "Mean", "N", "Median", "Mean", "Avg rounds")) |>
    kable(caption = "Review round duration, per submission year") |>
    add_header_above(c(
        " " = 1,
        "First round (days)" = 3,
        "Subsequent rounds (days)" = 3,
        " " = 1
    ))
```

::: {.callout-note collapse="true" title="Column definitions and caveats"}
- **First decision (days)** — days from submission to the first editorial
  decision on the paper.
- **Final decision (days)** — days from submission to the last decision, only
  for submissions that have reached a final status and went through at least
  one review round.
- **First round (days)** — round 1 duration: submission → round-1 closing
  decision.
- **Subsequent rounds (days)** — round N≥2 duration: previous round's close →
  this round's closing decision. All N≥2 rounds pooled.
- **Avg rounds** — mean number of review rounds per submission in the cohort.
- **Restriction**: submissions from 2024 onwards. Earlier papers were migrated
  from a previous OJS install and their decision history wasn't carried over.
:::

#### Round timing decomposed

Where editorial cycles actually spend their time: waiting on authors, finding
reviewers, waiting on reviews, or deciding.

```{r round-decomposition}
reviews_csv <- sort(Sys.glob("ojs_reports/reviews-*.csv"), decreasing = TRUE)[1]
reviews <- read.csv(reviews_csv, stringsAsFactors = FALSE, check.names = FALSE) |>
    rename(
        submission_id  = `Submission ID`,
        round          = Round,
        date_assigned  = `Date Assigned`,
        date_confirmed = `Date Confirmed`,
        date_completed = `Date Completed`,
        declined       = Declined,
        cancelled      = Cancelled
    ) |>
    mutate(
        date_confirmed = parse_dt(date_confirmed),
        date_completed = parse_dt(date_completed),
        completed_ok   = declined == "No" & cancelled == "No" &
                         !is.na(date_completed)
    )

per_round_reviews <- reviews |>
    group_by(submission_id, round) |>
    summarise(
        n_invited      = n(),
        n_completed    = sum(completed_ok),
        last_confirmed = suppressWarnings(max(date_confirmed[completed_ok],
                                              na.rm = TRUE)),
        last_completed = suppressWarnings(max(date_completed[completed_ok],
                                              na.rm = TRUE)),
        .groups        = "drop"
    ) |>
    mutate(
        last_confirmed = dplyr::if_else(is.finite(last_confirmed),
                                        last_confirmed, as.POSIXct(NA)),
        last_completed = dplyr::if_else(is.finite(last_completed),
                                        last_completed, as.POSIXct(NA))
    )

nrr_dates <- dec |>
    filter(label == "New Review Round") |>
    select(submission_id, nrr_date = date_decided) |>
    arrange(submission_id, nrr_date) |>
    group_by(submission_id) |>
    mutate(opens_round_num = row_number() + 1L) |>
    ungroup()

round_intervals_reviewed <- rounds |>
    select(submission_id, round_num, round_start, round_end) |>
    inner_join(per_round_reviews,
        by = c("submission_id", "round_num" = "round")
    ) |>
    left_join(
        nrr_dates |> select(submission_id,
                            round_num = opens_round_num, nrr_date),
        by = c("submission_id", "round_num")
    ) |>
    mutate(
        # Round 1 has no NRR (no revision phase); round N>=2 uses NRR as
        # the boundary. If NRR is missing for round N (unusual), fall back
        # to round_start.
        finding_start = dplyr::case_when(
            round_num == 1     ~ round_start,
            !is.na(nrr_date)   ~ nrr_date,
            TRUE               ~ round_start
        ),
        year          = substr(as.character(round_end), 1, 4),
        waiting_days  = ifelse(round_num == 1, NA_real_,
                               as.numeric(difftime(finding_start, round_start,
                                                   units = "days"))),
        finding_days  = as.numeric(difftime(last_confirmed, finding_start,
                                            units = "days")),
        review_days   = as.numeric(difftime(last_completed, last_confirmed,
                                            units = "days")),
        decide_days   = as.numeric(difftime(round_end, last_completed,
                                            units = "days")),
        total_days    = as.numeric(difftime(round_end, round_start,
                                            units = "days")),
        opened_date   = dplyr::case_when(
            round_num == 1   ~ as.POSIXct(NA),
            !is.na(nrr_date) ~ nrr_date,
            TRUE             ~ as.POSIXct(NA)
        ),
        kind          = "review"
    )

# Post-revision rounds: same review_round_id with both a revisions-style
# decision and a later Accept/Decline. Round_start = the revisions decision;
# round_end = the final decision; waiting = upload - rev_date; deciding =
# final - upload.
revision_labels <- c("Resubmit for Review", "Request Revisions")
final_labels    <- c("Accept Submission", "Decline Submission")

post_rev_events <- dec |>
    filter(!is.na(review_round_id), review_round_id > 0,
           label %in% c(revision_labels, final_labels)) |>
    group_by(submission_id, review_round_id) |>
    summarise(
        rev_date   = suppressWarnings(min(
            date_decided[label %in% revision_labels])),
        final_date = suppressWarnings(max(
            date_decided[label %in% final_labels])),
        .groups    = "drop"
    ) |>
    filter(is.finite(rev_date), is.finite(final_date),
           final_date > rev_date)

revision_files <- ojs_files_bulk(client,
    unique(post_rev_events$submission_id)) |>
    mutate(created_at = parse_dt(created_at)) |>
    filter(!is.na(created_at))

post_rev_intervals <- post_rev_events |>
    inner_join(revision_files, by = "submission_id",
               relationship = "many-to-many") |>
    filter(created_at >= rev_date, created_at <= final_date) |>
    group_by(submission_id, review_round_id, rev_date, final_date) |>
    summarise(upload_date = max(created_at), .groups = "drop") |>
    transmute(
        submission_id,
        round_num      = NA_integer_,
        round_start    = rev_date,
        round_end      = final_date,
        opened_date    = upload_date,
        n_invited      = 0L,
        n_completed    = 0L,
        last_confirmed = as.POSIXct(NA),
        last_completed = as.POSIXct(NA),
        year           = substr(as.character(final_date), 1, 4),
        waiting_days   = as.numeric(difftime(upload_date, rev_date,
                                             units = "days")),
        finding_days   = NA_real_,
        review_days    = NA_real_,
        decide_days    = as.numeric(difftime(final_date, upload_date,
                                             units = "days")),
        total_days     = as.numeric(difftime(final_date, rev_date,
                                             units = "days")),
        kind           = "post-revision"
    )

round_intervals <- bind_rows(round_intervals_reviewed, post_rev_intervals)

med <- function(x) round(median(x, na.rm = TRUE), 1)

interval_row <- function(df) {
    ok  <- df |> filter(kind == "post-revision" | n_completed > 0)
    nok <- df |> filter(kind == "review", n_completed == 0)
    tibble(
        N            = nrow(ok),
        N_post_rev   = sum(ok$kind == "post-revision"),
        waiting      = med(ok$waiting_days),
        finding      = med(ok$finding_days),
        reviewing    = med(ok$review_days),
        deciding     = med(ok$decide_days),
        total        = med(ok$total_days),
        no_reviews_N = nrow(nok)
    )
}

by_year_int <- round_intervals |>
    group_by(Year = year) |>
    group_modify(~ interval_row(.x)) |>
    ungroup() |>
    arrange(Year)

total_int <- interval_row(round_intervals) |>
    mutate(Year = "Total") |>
    select(Year, everything())

bind_rows(by_year_int, total_int) |>
    select(-N_post_rev, -no_reviews_N) |>
    setNames(c("Year", "N rounds",
               "Waiting", "Finding", "Reviewing", "Deciding",
               "Total round")) |>
    kable(caption = "Median days per phase of an editorial round (closing-decision year)")
```

::: {.callout-note collapse="true" title="Phase definitions and data sources"}
- **Waiting for revision** (round 2+ and post-revision rounds) — previous
  round's close → "New Review Round" decision (for review rounds) or revision
  file upload (for post-revision rounds).
- **Finding reviewers** — round start → last `Date Confirmed` among reviewers
  who ultimately completed a review.
- **Reviewing** — last `Date Confirmed` → last `Date Completed`.
- **Deciding** — last `Date Completed` (or revision upload for post-revision
  rounds) → round-closing decision.
- **Total round** — start of the round → its closing decision.

Two kinds of rounds are pooled:

1. **Review rounds** — author resubmits, editor sends out for review, reviews
   come in, editor decides.
2. **Post-revision rounds** — editor closed a previous round with
   `Request Revisions` / `Resubmit for Review` and then accepted/declined in
   the same `review_round_id` without opening another review round. No
   Finding/Reviewing, just Waiting + Deciding.

Review rounds where everyone declined or ghosted are excluded (shown
separately in the "w/o completed reviews" column). Year is the round's
closing-decision year.

Data: reviewer timeline from `ojs_reports/reviews-*.csv` (the API only exposes
deadlines, not actual confirm/complete timestamps); revision file uploads from
`/submissions/{id}/files?fileStages=15`.
:::

### Reviewer response and turnaround

Per-reviewer behavior on round-1 invitations: how many were invited, how many
finished a review, and how long each step took.

```{r reviewer-stats}
closed_r1 <- rounds |>
    filter(round_num == 1) |>
    pull(submission_id) |>
    unique()

reviewer_rows <- reviews |>
    filter(round == 1, submission_id %in% closed_r1) |>
    mutate(
        date_assigned = parse_dt(date_assigned),
        responded     = nzchar(as.character(date_confirmed)) &
                        !is.na(date_confirmed),
        accepted      = declined == "No"  & cancelled == "No" & responded,
        declined_resp = declined == "Yes" & cancelled == "No",
        ghosted       = !responded        & cancelled == "No",
        days_to_decline  = ifelse(declined_resp,
            as.numeric(difftime(date_confirmed, date_assigned, units = "days")),
            NA_real_),
        days_to_accept   = ifelse(accepted,
            as.numeric(difftime(date_confirmed, date_assigned, units = "days")),
            NA_real_),
        days_to_complete = ifelse(completed_ok,
            as.numeric(difftime(date_completed, date_confirmed, units = "days")),
            NA_real_)
    ) |>
    inner_join(
        subs_dt |> transmute(submission_id = id,
                             year = substr(as.character(date_submitted), 1, 4)),
        by = "submission_id"
    ) |>
    filter(year >= "2024")

mean_days <- function(x) {
    x <- x[!is.na(x)]
    if (length(x) == 0) "—" else sprintf("%.1f", mean(x))
}

reviewer_row <- function(df) {
    active <- df |> filter(cancelled == "No")
    n_subs <- length(unique(active$submission_id))
    tibble(
        invited     = ifelse(n_subs > 0,
            sprintf("%.1f", nrow(active) / n_subs), "—"),
        completed   = ifelse(n_subs > 0,
            sprintf("%.1f", sum(active$completed_ok) / n_subs), "—"),
        t_decline   = mean_days(active$days_to_decline),
        t_accept    = mean_days(active$days_to_accept),
        t_complete  = mean_days(active$days_to_complete),
        pct_accept  = ifelse(nrow(active) > 0,
            sprintf("%.0f%%", 100 * sum(active$accepted) / nrow(active)),
            "—"),
        pct_decline = ifelse(nrow(active) > 0,
            sprintf("%.0f%%", 100 * sum(active$declined_resp) / nrow(active)),
            "—"),
        pct_ghost   = ifelse(nrow(active) > 0,
            sprintf("%.0f%%", 100 * sum(active$ghosted) / nrow(active)),
            "—")
    )
}

by_year_rev <- reviewer_rows |>
    group_by(Year = year) |>
    group_modify(~ reviewer_row(.x)) |>
    ungroup() |>
    arrange(Year)

total_rev <- reviewer_row(reviewer_rows) |>
    mutate(Year = "Total") |>
    select(Year, everything())

bind_rows(by_year_rev, total_rev) |>
    setNames(c("Year", "Invited / ms.", "Completed / ms.",
               "Days to decline", "Days to accept", "Days to complete",
               "% accepted", "% declined", "% ghosted")) |>
    kable(caption = "Reviewer response and turnaround on round-1 invitations")
```

::: {.callout-note collapse="true" title="Column definitions"}
- **Invited / ms.** — average non-cancelled round-1 reviewer invitations per
  manuscript that received any.
- **Completed / ms.** — average reviewers per manuscript who accepted *and*
  submitted a review.
- **Days to decline / accept / complete** — per-reviewer means: time from
  invitation to decline response, invitation to acceptance, and acceptance to
  review submission.
- **% accepted / % declined / % ghosted** — share of non-cancelled
  invitations where the reviewer accepted, declined, or never responded. The
  three sum to 100%.

Restricted to the first review round (subsequent rounds typically re-invite the
same reviewers, which would double-count their behavior), to manuscripts
submitted from 2024 onward, and to manuscripts whose round 1 has closed with
an editorial decision (Accept / Decline / Request Revisions / Resubmit for
Review) — ongoing rounds are excluded so they don't pull down the completed
count. Year is the submission year.
:::

## AE Statistics

Current workload and recent throughput per active AE/EiC. Two tables: past
365 days, and all time.

```{r ae-performance}
aes <- read.csv("editors.csv", stringsAsFactors = FALSE) |>
    filter(role %in% c("AE", "EiC"), status == "current")

ae_subs <- purrr::map_dfr(aes$editor_id, function(eid) {
    ojs_submissions(client, assigned_to = eid) |>
        mutate(editor_id = eid)
})

cutoff <- as.POSIXct(Sys.Date() - 365, tz = "UTC")

first_dec <- decisions |>
    mutate(date_decided = parse_dt(date_decided)) |>
    filter(!is.na(date_decided)) |>
    group_by(submission_id) |>
    summarise(first_decision = min(date_decided), .groups = "drop")

# Final-decision date: last Accept/Decline on the paper.
final_dec <- decisions |>
    mutate(date_decided = parse_dt(date_decided)) |>
    filter(label %in% c("Accept Submission", "Decline Submission"),
           !is.na(date_decided)) |>
    group_by(submission_id) |>
    summarise(final_decision = max(date_decided), .groups = "drop")

ae_perf <- ae_subs |>
    mutate(
        date_submitted = parse_dt(date_submitted),
        outcome        = bucket(status_label)
    ) |>
    left_join(first_dec, by = c("id" = "submission_id")) |>
    mutate(days_to_first = as.numeric(
        difftime(first_decision, date_submitted, units = "days")
    ))

current_load <- ae_perf |>
    filter(outcome == "Pending") |>
    count(editor_id, name = "currently_assigned")

recent <- ae_perf |> filter(date_submitted >= cutoff)

total_summary <- recent |>
    group_by(editor_id) |>
    summarise(total_assigned = n(), .groups = "drop")

# First-decision time: only for first decisions taken in the past 365 days,
# regardless of submission date.
first_dec_summary <- ae_subs |>
    select(submission_id = id, editor_id,
           date_submitted_str = date_submitted) |>
    inner_join(first_dec, by = "submission_id") |>
    filter(first_decision >= cutoff) |>
    mutate(days_to_first = as.numeric(difftime(
        first_decision, parse_dt(date_submitted_str), units = "days"
    ))) |>
    group_by(editor_id) |>
    summarise(
        median_days_to_first = round(median(days_to_first, na.rm = TRUE), 1),
        .groups = "drop"
    )

# Final-decision time: same logic, but for the last Accept/Decline.
final_dec_summary <- ae_subs |>
    select(submission_id = id, editor_id,
           date_submitted_str = date_submitted) |>
    inner_join(final_dec, by = "submission_id") |>
    filter(final_decision >= cutoff) |>
    mutate(days_to_final = as.numeric(difftime(
        final_decision, parse_dt(date_submitted_str), units = "days"
    ))) |>
    group_by(editor_id) |>
    summarise(
        median_days_to_final = round(median(days_to_final, na.rm = TRUE), 1),
        .groups = "drop"
    )

# Average review rounds per paper that went through review (>=1 round).
# 365d window: papers whose *final decision* is in the past 365 days — same
# decision-date semantics as the other 365d metrics.
rounds_per_sub <- rounds |>
    group_by(submission_id) |>
    summarise(n_rounds = max(round_num), .groups = "drop")

avg_rounds_summary <- ae_subs |>
    select(submission_id = id, editor_id) |>
    inner_join(final_dec, by = "submission_id") |>
    filter(final_decision >= cutoff) |>
    inner_join(rounds_per_sub, by = "submission_id") |>
    group_by(editor_id) |>
    summarise(avg_rounds = round(mean(n_rounds), 2), .groups = "drop")

# Accept rate: any Accept/Decline decision on an AE's assigned paper, in the
# last 365 days. Counted by decision date, not submission date.
ae_decisions <- ae_subs |>
    select(submission_id = id, editor_id) |>
    inner_join(
        decisions |>
            mutate(date_decided = parse_dt(date_decided)) |>
            filter(!is.na(date_decided),
                   label %in% c("Accept Submission", "Decline Submission"),
                   date_decided >= cutoff) |>
            select(submission_id, label, date_decided),
        by = "submission_id",
        relationship = "many-to-many"
    )

accept_summary <- ae_decisions |>
    group_by(editor_id) |>
    summarise(
        accepted    = sum(label == "Accept Submission"),
        rejected    = sum(label == "Decline Submission"),
        accept_rate = ifelse(
            (accepted + rejected) > 0,
            round(100 * accepted / (accepted + rejected), 1),
            NA_real_
        ),
        .groups = "drop"
    )

aes |>
    select(editor_id, AE = full_name) |>
    left_join(current_load, by = "editor_id") |>
    left_join(total_summary, by = "editor_id") |>
    left_join(first_dec_summary, by = "editor_id") |>
    left_join(final_dec_summary, by = "editor_id") |>
    left_join(avg_rounds_summary, by = "editor_id") |>
    left_join(accept_summary, by = "editor_id") |>
    mutate(
        currently_assigned = tidyr::replace_na(currently_assigned, 0L),
        total_assigned     = tidyr::replace_na(total_assigned, 0L),
        accepted           = tidyr::replace_na(accepted, 0L),
        rejected           = tidyr::replace_na(rejected, 0L),
        decided            = accepted + rejected,
        accept_rate_str    = ifelse(
            decided == 0,
            "—",
            sprintf("%.0f%% (%d/%d)", accept_rate, accepted, decided)
        )
    ) |>
    select(
        AE,
        `Currently assigned`     = currently_assigned,
        `Assigned (365d)`        = total_assigned,
        `Days to 1st decision`   = median_days_to_first,
        `Days to final decision` = median_days_to_final,
        `Avg rounds`             = avg_rounds,
        `Accept rate (decided)`  = accept_rate_str
    ) |>
    kable(caption = sprintf(
        "AE statistics past 365 days (since %s)",
        format(Sys.Date() - 365)
    ))
```

```{r ae-performance-alltime}
# All-time: total assigned, median time-to-first-decision, accept rate.
total_summary_all <- ae_subs |>
    group_by(editor_id) |>
    summarise(total_assigned = n(), .groups = "drop")

first_dec_summary_all <- ae_subs |>
    select(submission_id = id, editor_id,
           date_submitted_str = date_submitted) |>
    inner_join(first_dec, by = "submission_id") |>
    mutate(days_to_first = as.numeric(difftime(
        first_decision, parse_dt(date_submitted_str), units = "days"
    ))) |>
    group_by(editor_id) |>
    summarise(
        median_days_to_first = round(median(days_to_first, na.rm = TRUE), 1),
        .groups = "drop"
    )

final_dec_summary_all <- ae_subs |>
    select(submission_id = id, editor_id,
           date_submitted_str = date_submitted) |>
    inner_join(final_dec, by = "submission_id") |>
    mutate(days_to_final = as.numeric(difftime(
        final_decision, parse_dt(date_submitted_str), units = "days"
    ))) |>
    group_by(editor_id) |>
    summarise(
        median_days_to_final = round(median(days_to_final, na.rm = TRUE), 1),
        .groups = "drop"
    )

avg_rounds_summary_all <- ae_subs |>
    select(submission_id = id, editor_id) |>
    inner_join(rounds_per_sub, by = "submission_id") |>
    group_by(editor_id) |>
    summarise(avg_rounds = round(mean(n_rounds), 2), .groups = "drop")

accept_summary_all <- ae_subs |>
    select(submission_id = id, editor_id) |>
    inner_join(
        decisions |>
            filter(label %in% c("Accept Submission", "Decline Submission")) |>
            select(submission_id, label),
        by = "submission_id",
        relationship = "many-to-many"
    ) |>
    group_by(editor_id) |>
    summarise(
        accepted    = sum(label == "Accept Submission"),
        rejected    = sum(label == "Decline Submission"),
        accept_rate = ifelse(
            (accepted + rejected) > 0,
            round(100 * accepted / (accepted + rejected), 1),
            NA_real_
        ),
        .groups = "drop"
    )

aes |>
    select(editor_id, AE = full_name) |>
    left_join(current_load, by = "editor_id") |>
    left_join(total_summary_all, by = "editor_id") |>
    left_join(first_dec_summary_all, by = "editor_id") |>
    left_join(final_dec_summary_all, by = "editor_id") |>
    left_join(avg_rounds_summary_all, by = "editor_id") |>
    left_join(accept_summary_all, by = "editor_id") |>
    mutate(
        currently_assigned = tidyr::replace_na(currently_assigned, 0L),
        total_assigned     = tidyr::replace_na(total_assigned, 0L),
        accepted           = tidyr::replace_na(accepted, 0L),
        rejected           = tidyr::replace_na(rejected, 0L),
        decided            = accepted + rejected,
        accept_rate_str    = ifelse(
            decided == 0,
            "—",
            sprintf("%.0f%% (%d/%d)", accept_rate, accepted, decided)
        )
    ) |>
    select(
        AE,
        `Currently assigned`     = currently_assigned,
        `Assigned (all time)`    = total_assigned,
        `Days to 1st decision`   = median_days_to_first,
        `Days to final decision` = median_days_to_final,
        `Avg rounds`             = avg_rounds,
        `Accept rate (decided)`  = accept_rate_str
    ) |>
    kable(caption = "AE statistics all time")
```

::: {.callout-note collapse="true" title="Column definitions"}
- **Currently assigned** — submissions where the AE is the responsible editor
  (OJS `assignedTo`) and the submission is still in flight (not declined,
  published, or in production). No date restriction.
- **Assigned (365d / all time)** — submissions assigned to the AE whose
  `dateSubmitted` falls within the window.
- **Days to 1st decision** — for each first decision *taken* in the window on
  the AE's assigned papers, days from submission to that decision, median
  across those decisions.
- **Days to final decision** — same logic but for the last `Accept Submission`
  or `Decline Submission` decision on the paper.
- **Avg rounds** — average number of review rounds per paper that went through
  review (papers with no review rounds excluded). For the 365d table, restricted
  to papers whose final decision is in the past 365 days.
- **Accept rate (decided)** — of all `Accept Submission` / `Decline Submission`
  decisions on the AE's assigned papers, taken in the window, percent
  accepted. Numerator/denominator shown in parentheses.

The 365-day window cuts by *decision date* for the time-to-decision medians
and the accept rate, but by *submission date* for the "Assigned (365d)" count.
:::

#### Round timing per AE

Same intervals as the decomposed table above, broken down by handling AE/EiC.
Small N — single slow rounds will move medians a lot.

```{r ae-round-timing}
ae_round_summary_fn <- function(intervals) {
    intervals |>
        group_by(editor_id) |>
        summarise(
            n_rounds   = n(),
            waiting    = med(waiting_days),
            finding    = med(finding_days),
            reviewing  = med(review_days),
            deciding   = med(decide_days),
            total      = med(total_days),
            .groups    = "drop"
        )
}

ae_round_render <- function(summary_df, caption) {
    aes |>
        select(editor_id, AE = full_name) |>
        left_join(summary_df, by = "editor_id") |>
        mutate(n_rounds = tidyr::replace_na(n_rounds, 0L)) |>
        select(
            AE,
            `N rounds`    = n_rounds,
            Waiting       = waiting,
            Finding       = finding,
            Reviewing     = reviewing,
            Deciding      = deciding,
            `Total round` = total
        ) |>
        kable(caption = caption)
}

ae_rounds_365 <- ae_subs |>
    select(submission_id = id, editor_id) |>
    inner_join(
        round_intervals |>
            filter(kind == "post-revision" | n_completed > 0,
                   round_end >= cutoff),
        by = "submission_id",
        relationship = "many-to-many"
    )

ae_round_render(
    ae_round_summary_fn(ae_rounds_365),
    sprintf("Median days per round phase, per AE (rounds closed since %s)",
            format(Sys.Date() - 365))
)
```

```{r ae-round-timing-alltime}
ae_rounds_all <- ae_subs |>
    select(submission_id = id, editor_id) |>
    inner_join(
        round_intervals |>
            filter(kind == "post-revision" | n_completed > 0),
        by = "submission_id",
        relationship = "many-to-many"
    )

ae_round_render(
    ae_round_summary_fn(ae_rounds_all),
    "Median days per round phase, per AE (all time)"
)
```

<style>
.dt-raw .dataTables_wrapper,
.dt-raw table.dataTable,
.dt-raw table.dataTable th,
.dt-raw table.dataTable td {
    font-size: 12px;
}
.dt-raw table.dataTable th,
.dt-raw table.dataTable td {
    padding: 3px 6px;
}
/* Shrink the per-column filter inputs in the header */
.dt-raw .dataTables_wrapper input[type="search"],
.dt-raw .dataTables_wrapper .form-control {
    font-size: 12px;
    height: auto;
    padding: 1px 22px 1px 6px;
    line-height: 1.3;
}
.dt-raw .dataTables_wrapper .form-group {
    margin: 0;
}
.dt-raw .dataTables_wrapper .form-control-feedback {
    height: auto;
    line-height: 1.6;
    font-size: 11px;
}
</style>

#### Detailed information per review round

Per-round raw timestamps for every review and post-revision round. Sortable
by clicking any column header (date columns sort by elapsed days, not text).
Cells shaded green → red by the displayed day count.

::: {.dt-raw}

```{r ae-round-raw}
fmt_gap <- function(date, prior = NULL) {
    months <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
    base <- ifelse(
        is.na(date), "—",
        sprintf("%d %s",
                as.integer(format(date, "%d")),
                months[as.integer(format(date, "%m"))])
    )
    if (is.null(prior)) return(base)
    days <- as.integer(round(as.numeric(
        difftime(date, prior, units = "days"))))
    ifelse(!is.na(date) & !is.na(prior),
           sprintf("%s (%dd)", base, days), base)
}

ae_subs |>
    select(submission_id = id, editor_id) |>
    inner_join(
        round_intervals |>
            filter(kind == "post-revision" | n_completed > 0),
        by = "submission_id",
        relationship = "many-to-many"
    ) |>
    inner_join(aes |> select(editor_id, AE = full_name), by = "editor_id") |>
    mutate(AE = sub(" .*", "", AE)) |>
    arrange(submission_id, round_end) |>
    group_by(submission_id) |>
    mutate(display_round = row_number()) |>
    ungroup() |>
    arrange(AE, submission_id, display_round) |>
    mutate(
        upload_actual   = coalesce(opened_date, round_start),
        request_display = dplyr::if_else(
            !is.na(round_num) & round_num == 1,
            as.POSIXct(NA), round_start
        ),
        decided_prev    = coalesce(last_completed, last_confirmed,
                                   upload_actual)
    ) |>
    transmute(
        AE,
        Year              = substr(as.character(upload_actual), 1, 4),
        `Subm.`           = submission_id,
        Rnd               = display_round,
        Requested         = fmt_gap(request_display),
        Uploaded          = fmt_gap(upload_actual, request_display),
        `Reviewers found` = fmt_gap(last_confirmed, upload_actual),
        `Reviews done`    = fmt_gap(last_completed, last_confirmed),
        Decided           = fmt_gap(round_end, decided_prev),
        waiting_d   = as.integer(round(as.numeric(
            difftime(upload_actual, request_display, units = "days")))),
        finding_d   = as.integer(round(as.numeric(
            difftime(last_confirmed, upload_actual, units = "days")))),
        reviewing_d = as.integer(round(as.numeric(
            difftime(last_completed, last_confirmed, units = "days")))),
        deciding_d  = as.integer(round(as.numeric(
            difftime(round_end, decided_prev, units = "days")))),
        requested_n = as.numeric(request_display)
    ) |>
    datatable(
        rownames = FALSE,
        filter   = "top",
        class    = "compact stripe hover row-border",
        options  = list(
            pageLength = 20,
            scrollX    = TRUE,
            columnDefs = list(
                list(visible = FALSE, targets = c(9, 10, 11, 12, 13)),
                list(orderData = 13, targets = 4),  # Requested → ts
                list(orderData = 9,  targets = 5),  # Uploaded → waiting
                list(orderData = 10, targets = 6),  # Reviewers found
                list(orderData = 11, targets = 7),  # Reviews done
                list(orderData = 12, targets = 8)   # Decided → deciding
            )
        )
    ) |>
    formatStyle("Uploaded", "waiting_d",
        backgroundColor = styleInterval(
            c(20, 40, 60, 80, 100),
            c("#c8e6c9", "#dcedc8", "#fff9c4",
              "#ffe0b2", "#ffcdd2", "#ef9a9a"))) |>
    formatStyle("Reviewers found", "finding_d",
        backgroundColor = styleInterval(
            c(20, 40, 60, 80, 100),
            c("#c8e6c9", "#dcedc8", "#fff9c4",
              "#ffe0b2", "#ffcdd2", "#ef9a9a"))) |>
    formatStyle("Reviews done", "reviewing_d",
        backgroundColor = styleInterval(
            c(20, 40, 60, 80, 100),
            c("#c8e6c9", "#dcedc8", "#fff9c4",
              "#ffe0b2", "#ffcdd2", "#ef9a9a"))) |>
    formatStyle("Decided", "deciding_d",
        backgroundColor = styleInterval(
            c(20, 40, 60, 80, 100),
            c("#c8e6c9", "#dcedc8", "#fff9c4",
              "#ffe0b2", "#ffcdd2", "#ef9a9a")))
```

:::

::: {.callout-note collapse="true" title="Column definitions and sorting"}
- **Year** — year of the Uploaded timestamp.
- **Subm.** — OJS submission ID.
- **Rnd** — round number per submission, in chronological order; post-revision
  rounds get the next number after the last review round.
- **Requested** — editor's revision-request decision (Resubmit / Request
  Revisions) that opened this round. Empty for round 1.
- **Uploaded** — when the author handed something over: original submission
  date for round 1, "New Review Round" click for review rounds 2+, revision
  file upload date for post-revision rounds.
- **Reviewers found** — last `Date Confirmed` among reviewers who completed
  a review (empty for post-revision rounds — no reviewers).
- **Reviews done** — last `Date Completed` among reviewers (empty for
  post-revision rounds).
- **Decided** — round-closing decision date.

Each non-leftmost date cell shows the gap from the previous milestone in the
row (e.g. `8 Apr (12d)`). Clicking a date column sorts by that gap (or by
chronology for Requested), not alphabetically. Cell background is shaded by
day count: green ≤20, yellow ~50, red >100.
:::

:::

